\cleardoublepage
\phantomsection
\pdfbookmark{Acknowledgements}{Acknowledgements}

\begin{flushright}{
    \slshape
    ``We don't read and write poetry because it's cute. We read and write poetry because we are members of the human race. And the human race is filled with passion. And medicine, law, business, engineering, these are noble pursuits and necessary to sustain life. But poetry, beauty, romance, love, these are what we stay alive for.''} \\
    \medskip
    --- N.H. Kleinbaum, Dead Poets Society
\end{flushright}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}

\noindent First and foremost, I would like to express my gratitude to Prof. Gaggi, given her support throughout two paths of thesis, both in bachelor and master degrees, for valuable knowledge and support throughout these academic years, both humanly and academically. 

\vspace{0.35cm}

\noindent I would like to thank my mom, the only person who supported me practically throughout these years and gave me many life lessons, maybe not in the right way, but her love was always present for me. For the same reason, my life has been very dense of things, but I promised myself I would have always been able to make it. And I did it always on my own, in ways I would have never imagined since I had no guidance, but I was always everyone guidance. Not arrogance, simply human nature. I hope after this huge step to conquer the long awaited peace of spirit and soul. 

\vspace{0.35cm}

\noindent A special thank you to the few real friends I have (particularly Matilde and Andrea), because they helped me do many things up until now and I would not live a day without them.

\vspace{0.75cm}

\noindent{\myLocation, \myTime}
\hfill \textit{\myName}

\endgroup
